---
title: "Sales case"
author: "CRJR"
output: html_document
---

```{r setup, include=FALSE}

library(tidyverse)
library(reshape2)
library(ggplot2)
library(plotly)
library(rpivotTable)
library(keras)
library(lubridate)
library(dplyr)
library(forecast)
library(xts)
library(data.table)

setwd("/Users/Cristhian/Documents/Sales/")
data <- read.csv("Sales.csv", header = TRUE, sep = ",")

dictDates <- as.data.frame(ymd("2016-12-01") + days(0:395))
colnames(dictDates)<-c('date')

data$date <- as.Date(data$date)
data <- left_join(dictDates,data)

data$sales<- ifelse(is.na(data$sales),0,data$sales)

initialSummary <- data %>%
  group_by(country) %>%
  summarise(count = n_distinct(customer),
            sumSales = sum(sales)) %>%
  mutate(proportionSales = 100*sumSales/sum(sumSales),
         proportionCountries = 100*count/sum(count))
```

This RMD is divided in:

* Data discovery
* Data transformation and clusters
* Sales forecasting 

This data contains sales information from a company. The aim of this exercise is to understand it and to provide a forecast for december 2017.

#Data discovery
```{r}
first(as.Date(data$date))
last(as.Date(data$date))

data <- data %>%
  filter(date<=as.Date('2017-12-01'))

initialVisualization <- data %>%
  group_by(date) %>%
  summarise(sumSales = sum(sales))

ggplot(initialVisualization)+
  geom_line(aes(x=date,y=sumSales, group=1))+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


To better visualize and understand the data, we will change the timestamp and aggregate the sales column.

* Change the timestamp and aggregate the sales 

```{r}
Visualization1 <- data %>%
  group_by(date = cut(as.Date(date) ,'week'))%>%
  summarise(sumSales = sum(sales))

ggplot(Visualization1)+
  geom_line(aes(x=date,y=sumSales, group=1))+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Some basic numbers to understand the data:
* Number of different customers = 4,372
* Number of different countries = 37 (including Unspecified)
* Sales = 8,300,065.81 Units / Euro ?
* Number of different products = 3,684 
* Average number of sales per customer = 93.05329
* Average sales per customer = 28.83901, but standard deviation = 


```{r, include=FALSE}
Visualization2 <- data %>%
  group_by(country)%>%
  summarise(customer = n_distinct(customer))
table(Visualization2)

Visualization3 <- data %>%
  group_by(customer)%>%
  summarise(sales = sum(sales),
            n_sales = n()) %>%
  mutate(avgSale = sales/n_sales )

Visualization4 <- data %>%
  group_by(date = cut(as.Date(date) ,'week'))%>%
  summarise(sales = sum(sales),
            n_sales = n()) %>%
  mutate(avgSale = sales/n_sales )
```

The previous chart suggests that we could apply a difference in order to eliminate the trend on the data over time. The ACF and PACF charts will confirm that.
```{r}
acf(Visualization1$sumSales)
pacf(Visualization1$sumSales)
```

Before going further and choosing the P and Q parameters of the ARIMA approach, we need to find out if it makes sense to model the average sale over the time. The following will show if this makes sense.
```{r}
ggplot(Visualization4)+
  geom_line(aes(x=date,y=avgSale, group=1))+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

So far, it seems that by performing transformations to the data we could model with ARIMA models the sales information. Before that, let us see if all the customers are similar in terms of amount of information, sales, and average sales.

#Data transformation and clusters
```{r}
GroupSales <- data %>%
  group_by(customer)%>%
  summarise(sales = sum(sales), count = n())%>%
  mutate(limitSales = ifelse((sales-mean(sales))/sd(sales)<=2.5, 1, 0)) 

hist(GroupSales$sales,breaks = 100)

km <- kmeans(c(
  (GroupSales$sales - mean(GroupSales$sales))/sd(GroupSales$sales), 
  (GroupSales$count - mean(GroupSales$count))/sd(GroupSales$count) ),5)


```

the previous shows that:
* 34 customers makes ~27% of the sales
* 4338 customers makes the other ~73% of the sales

Where are we now? What next?

* Creating a unique model for the whole set of customers could lead to a wrong generalization and to a wrong forecast.
* We will create groups of clientes based on their sales levels and number of sales.

Firstly, we will split the customers in two, the ones that produce 27% of the sales and the ones that produce 73% of the sales.

```{r}

GroupSales1 <- GroupSales %>%
  filter(limitSales==1)  

GroupSales0 <- GroupSales %>%
  filter(limitSales==0)  

ggplot(GroupSales1)+
  geom_point(aes( x=(sales-mean(sales))/sd(sales), y=(count-mean(count))/sd(count)))+
  theme_minimal()

ggplot(GroupSales0)+
  geom_point(aes( x=(sales-mean(sales))/sd(sales), y=(count-mean(count))/sd(count)))+
  theme_minimal()

```

Futhermore, we will use kmeans for creating clusters of customers and then create models to forecast the sales. This is how the clusters look like:

```{r}
gs1<- select(GroupSales1,sales,count)
gs1$sales<-(gs1$sales-mean(gs1$sales))/ sd(gs1$sales)
gs1$count<-(gs1$count-mean(gs1$count))/ sd(gs1$count)
km1 <- kmeans(gs1,5)
ggplot(GroupSales1, aes( x=sales, y=count, color=km1$cluster))+
  geom_point()+
  theme_minimal()


GroupSales1<- GroupSales1 %>% add_column(cluster = km1$cluster)
gs1Dict <- GroupSales1 %>% select(customer,cluster)

custDict <- data %>% 
  group_by(customer) %>%
  summarise(count = n())%>%
  select(customer)

custDict <- left_join( custDict , gs1Dict)
custDict$cluster <- ifelse(is.na(custDict$cluster),0,custDict$cluster)
```

#Sales forecasting 

* Build an ARIMA model for every group of clients.
* Forecast sales of December 2017 based on the model from where the customer belongs to.
```{r}
ModelsByClst <- function(custDict, data){
  clModels <- list()
  for(i in 0:5){
    toTs <- custDict %>%
      filter(cluster == i) %>%
      left_join(data) %>%
      group_by(date = cut(as.Date(date) ,'week')) %>%
      summarise(sales = sum(sales), count_clients = n_distinct(customer) )%>%
      mutate(avgSales = sales/count_clients )%>%
      filter(as.Date(date) < as.Date('2017-12-01'))
    
    clTs <- xts(toTs$avgSales,as.Date(toTs$date))  
    fit <- auto.arima(clTs)
    print(i)
    print(fit)
    clModels[[i+1]] <- fit
    }
  return(clModels)
}

mods <- ModelsByClst(custDict ,data)
dictDates <- as.data.frame(ymd("2016-12-01") + weeks(0:57))
colnames(dictDates)<-c("date")


ForDecSls <- function(custDict){
  CustList <- list()
  SlsList <- list()
  compData <- data.frame(customer = integer(), sales= integer())
  for (i in 1:length(custDict$customer)){
    nCustomer <- custDict$customer[i]
    nCluster <- custDict$cluster[i] +1
    
    toTs <- data %>%
      filter(customer == nCustomer) %>%
      group_by(date) %>%
      summarise(sales=sum(sales))
    
    toTs$date<-as.Date(as.character(toTs$date))
    
    toTs <- left_join(dictDates,toTs)
    
    toTs$sales <- ifelse(is.na(toTs$sales),0,toTs$sales)
    
    toTs <- toTs %>%
      group_by(date = cut(as.Date(date) ,'week')) %>%
      summarise(sales = sum(sales))%>%  
      filter(as.Date(date) < as.Date('2017-12-01'))
    
    clTs <- xts(toTs$sales,as.Date(toTs$date))  
    fit<-Arima(clTs,model = mods[[nCluster]])
    forSales<-forecast(fit, h=4)
    slsDec <- sum(forSales[4]$mean[1:4])
    
    compData <- rbind(compData,c(nCustomer, slsDec))
    
  }
  
  colnames(compData) <- c('Customer','SlsDecember')
  return(compData)
  
}
```

The ARIMA models and coefficients vary significantly from every group of customes, this could mean that we have created a good segmentation because, on the one side, the customers on every single cluster have similar amount of information and sales's level, and on the other hand, the ARIMA models vary.

Finally we can compute the predicted sales for December 2017
```{r, include= FALSE}
DecemberSls <- ForDecSls(custDict)
```

```{r}
View(DecemberSls)
```

